---
title: "Intro to R Part 2 Notes"
output: html_document
date: "2023-09-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# First step: require tidyverse

```{r}
require(tidyverse)
```

# Loading data

We can load data either by downloading it to our computer, or just loading it directly from the internet with the function `read_rds()`.

```{r}
df <- read_rds("https://github.com/jbisbee1/DS1000_F2023/raw/main/Lectures/2_Intro_to_R/data/sc_debt.Rds")

df
```

# Functions for looking at data

The first step of the scientific method is **OBSERVATION**

```{r}
df %>% # Pipes or "do operators"
  select(instnm,ugds,costt4_a,control)

# Looking at column names
colnames(df)

# Using filter() to look at ROWS
df %>%
  filter(instnm == "Vanderbilt University") %>%
  select(ugds,costt4_a,sat_avg,md_earn_wne_p6)
```

# Comparing us to others

We want to know where we stand relative to other schools.

```{r}
df %>%
  arrange(sat_avg) %>% # Default is lowest to highest
  select(instnm,sat_avg)

df %>%
  arrange(desc(sat_avg)) %>% # Use desc() to flip sorting
  select(instnm,sat_avg) %>%
  print(n = 20)
```

# Creating a new variable

We use the function `mutate()` to create new variables.

```{r}
df %>%
  mutate(new_variable = 1) %>%
  select(unitid,new_variable)

df %>%
  mutate(sel = ifelse(test = adm_rate < .1,
                      yes = 1,
                      no = 0)) %>%
  arrange(adm_rate) %>%
  select(instnm,adm_rate,sel) %>%
  print(n = 20)

# Assign the new column with <- 
df <- df %>%
  mutate(sel = ifelse(adm_rate < .1,1,0))
```

# Compare with a filter

```{r}
df %>%
  filter(sel == 1) %>%
  arrange(desc(sat_avg)) %>%
  select(instnm,sat_avg)
```