---
title: "Intro to R Part 3 Notes"
output: html_document
date: "2023-09-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting started

```{r}
require(tidyverse)

df <- read_rds('https://github.com/jbisbee1/DS1000_F2023/raw/main/Lectures/2_Intro_to_R/data/sc_debt.Rds')
```

# Taking averages

- `mean()`: function for taking the average of some numbers

```{r}
# Example with mutate()
df %>%
  mutate(avg_avg_sat = mean(sat_avg,na.rm = TRUE)) %>%
  select(avg_avg_sat,sat_avg)

# Example using summarise() / summarize()
df %>%
  summarise(avg_avg_sat = mean(sat_avg,na.rm=T))

# Average admissions rate
df %>%
  summarise(adm_avg = mean(adm_rate,na.rm=T))
```

# Using filter() to look at different schools

- Use `filter()` to first calculate average SAT for selective schools

```{r}
df %>%
  filter(adm_rate < .1) %>%
  summarise(avg_sat_selective = mean(sat_avg,na.rm=T))

df %>%
  filter(adm_rate >= .1) %>%
  summarise(avg_sat_nonselective = mean(sat_avg,na.rm=T))
```

# Introducing the `group_by()` function

```{r}
df %>%
  mutate(sel = ifelse(adm_rate <.1,1,0)) %>%
  group_by(sel) %>%
  summarise(sat_avg_RQanswer = mean(sat_avg,na.rm=T))
```

# More fine-grained "selective" schools

```{r}
df <- df %>%
  mutate(sel_3levels = ifelse(adm_rate > .5,"Not Selective",
                              ifelse(adm_rate > .2,"Somewhat Selective","Very Selective")))

df %>%
  select(adm_rate,selective,sel_3levels)

df %>%
  group_by(sel_3levels) %>%
  summarise(sat_avg = mean(sat_avg,na.rm=T))


```

# Visualizing in `R`

```{r}
df %>%
  ggplot(aes(x = adm_rate,y = sat_avg,label = instnm)) +
  geom_point() +
  geom_smooth(method = 'lm') + 
  geom_text(size = 2)
```

# Other types of plots

- What region has the highest average SAT score? What about the lowest?

```{r}
df %>%
  group_by(region) %>%
  summarise(avg_sat = mean(sat_avg,na.rm=T)) %>%
  arrange(avg_sat) %>%
  ggplot(aes(y = reorder(region,avg_sat),x = avg_sat)) + # Trick #1: reorder the axes
  geom_bar(stat = 'identity')

df %>%
  ggplot(aes(x = region)) + 
  geom_bar()
```

